---
title: "Analysis of Ecosystems homework week 5"
subtitle: "Intro to statistical inference"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "The assignment"
output: pdf_document
---

```{r setup, echo=FALSE, results='hide'}
# Code Chunk 1
knitr::opts_chunk$set(message = FALSE, warning=FALSE, echo=FALSE, eval=FALSE)
```

```{r packages, results='hide'}
# Code Chunk 2
if (!require("pacman")) install.packages("pacman")
pacman::p_load(s20x, plyr, ggplot2, car, gvlma)
```

# Data preparation 

This assignment will use data in the \texttt{mpg} data from \texttt{ggplot2}\textemdash \texttt{cty} and \texttt{displ} are provided, and script to add the predictor variable, \texttt{origin}, is already in Code Chunk 3 for you. Note that \texttt{echo=FALSE} so you should not
see this script in the Word file, but \texttt{results=’verbatim’} so we should see the results of the call to \texttt{str()}.

```{r results='verbatim'}
# R code chunk 3: Add columns
# mpg$auto.trans <- with(mpg, ifelse((substring(trans,1,4)=='auto'),1, 0))
mpg$origin <- as.factor(revalue(mpg$manufacturer, 
                                c("audi"="Foreign","chevrolet"="USA",
                                  "dodge"="USA", "ford"="USA","jeep"="USA", 
                                  "honda"="Foreign", "hyundai"="Foreign",
                                  "land rover"="Foreign","subaru"="Foreign", 
                                  "lincoln"="USA", "mercury"="USA", 
                                  "nissan"="Foreign", "pontiac"="USA", 
                                  "toyota"="Foreign","volkswagen"="Foreign"))) 
str(mpg)
```

# The t test

## Assumptions

Briefly describe three assumptions of the t test. Recall that two of them relate to the data themselves and a third is more about experimental design.

## Distribution 

### Graphing

Make one or more graphs that visualizes the distribution of the response variable \texttt{mpg\$cty} with respect to the assumptions of the t test model. Use and copy Code Chunk 4 as necessary. Relevant items in your toolbox include histograms and density plots, faceting and colour aesthetics, Q-Q plots, and transformations. 


```{r results='verbatim', fig=TRUE}
# Code Chunk 4: distribution assessment 
  # Add additional code chunks as necessary.




```

### Interpretation

Below the code chunk, respond to the following:

* What do you conclude about the distribution of these data with respect to the assumptions of the t test?
* Describe what modifications, if any, you applied to the data to better fit model assumptions.

## Variance

### Test

In Code Chunk 5, conduct one or more relevant statistical tests to assess whether variance in the response variable \texttt{mpg\$cty} meets the assumptions of the t test model. Manipulate the data as necessary.


```{r results='verbatim'}
# Code Chunk 5: variance assessment 
  # Add additional code chunks as necessary



```

### Interpretation

Below the code chunk, respond to the following:

* What do you conclude about the variance of these data with respect to the assumptions of the t test?
* Describe the modification you applied to the data to better fit model assumptions.

# The linear model

## Assumptions

Briefly describe four assumptions of the linear regression model.

## Distribution 

### Graphing

Make one or more graphs that visualize the distribution of the response variable \texttt{mpg\$cty} with respect to the assumptions of the linear model. Use and copy Code Chunk 6 as necessary. Relevant items in your toolbox include histograms and density plots, Q-Q plots, and transformations.


```{r results='verbatim', fig=TRUE}
# Code Chunk 6: distribution assessment 
  # Add additional code chunks as necessary.



   
```

### Interpretation

Below the code chunk, respond to the following:

* What do you conclude about the distribution of these data with respect to the assumptions of the linear model?
* Describe the modification you applied to the data to better fit model assumptions.

## Fit a linear model

### Test

In Code Chunk 7, use \texttt{cty} and \texttt{displ} from \texttt{mpg} as response and predictor variables, respectively, to:

* Fit a linear model using formula notation
* Provide summary statistics of the model


```{r  results='verbatim'}
# Code Chunk 7: Define a linear model



```

### Plot

In Code Chunk 8, produce an appropriate graph of the linear model fit above using \texttt{ggplot}. Include the following:

* Data
* Trendline
* Informative axis labels


```{r results='verbatim', fig=TRUE}
# Code Chunk 8: Graph two continuous variables and fit a regression line




```

## Evaluate the linear model

### Model fit

Use Code Chunk 9 to evaluate the model fit with respect to assumptions of the linear model. Below the Code Chunk:

* Give an interpretation of model fit
* Describe what you base this assessment on


```{r  results='verbatim', fig=TRUE}
# Code Chunk 9: Q-Q Plot of residuals and test of model assumptions


```


### Model results

Report the following from the linear model (enter as text, not \texttt{R} output). When math is required on your part using information from R output, show your work. Provide both the letter that designates the test statistic/parameter as well as the value, separated by an = sign :

* Degrees of freedom
* Total Sum of Squares (show your math)
* Test statistic for the overall model
* Test statistic for the dependent variable
* Results of significance test
* How much variation is explained by the linear model?

### Interpretation

Below the code chunk, give your conclusion about the relationship between these variables.

