---
title: "Principal Components Analysis"
subtitle: "Homework Week 10"
author: "The Solution"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output: pdf_document
header-includes:
  - \usepackage{subfig}
---

```{r eval=TRUE, message=FALSE, echo=FALSE}
# Code Chunk 1
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=TRUE, eval=TRUE, fig.align = 'center')
if (!require("pacman")) install.packages("pacman")
source('C:/Users/Devan.McGranahan/GoogleDrive/Computer stuff/code snippets/ggplot themes/theme_ord.R')
```

```{r }
pacman::p_load(vegan, tidyverse, broom, ggordiplots, RVAideMemoire, pander )
```

# Data preparation, identification


```{r }
 load("C:/Users/Devan.McGranahan/GoogleDrive/Teaching/Classes/Analysis of Ecosystems/compiled notes/AoE course materials/example datasets/mtcars2.Rdata")
str(mtcars2)

```

\begin{itemize}
\item Which variables look appropriate for inclusion in the site x species matrix (ordination data)?
\item[] \textsf{ The continuous variables \texttt{disp}\textendash \texttt{mpg} are good variables for a PCA. \\}

\item Which might be “environmental” variables to test?
\item[] \textsf{Any of the multi-level factor variables that have several items/level; \texttt{make.model} and \texttt{make} do not have enough levels to test variabilty.  \\}
\end{itemize}

\clearpage

# Analysis 

## Fit and assess ordination 

### Fit the PCA


```{r results='hide'}
cars.d <- 
  mtcars2 %>%
    select(., .data$disp :.data$mpg )
 
cars.pca <- rda(cars.d ~ 1, scale=TRUE)
```

### Assess the PCA

```{r results='hide', fig.cap='Two approaches to making a screeplot for the \\texttt{mtcars2} PCA. \\label{scree}', fig.subcap=c('Plain old \\texttt{base} graphics.', '\\texttt{ggplot} can handle ordination objects when \\texttt{tibble::enframe} is applied to the correct part of the object (\\texttt{eig)}.'), fig.width=3, fig.asp=1, fig.ncol = 2}
par(mar=c(4,4,1,1)) 
screeplot(cars.pca, type="l", main=" ")
  
enframe(cars.pca$CA$eig) %>%
  plyr::rename(c("value"="Inertia")) %>%
ggplot() + theme_bw(14) + 
  geom_path(aes(x=name, y=Inertia, group=1), 
            color="darkblue", size=1.5) + 
  labs(x="Principal component axis")
```

\textsf{The screeplot (Fig. \ref{scree}) suggests marginal reductions in inertia (total variation explained) beyond PC2. \\ }

```{r assessment}
round(summary(eigenvals(cars.pca)), 2 ) %>%
  pander(caption="Eigenvalues and proportion explained by each axis in the PCA.")
round(scores(cars.pca, choices=c(1:6))$species, 2) %>%
  pander("Loadings for each variable along each PC, representing the relative influence 
         of each variable on variation in each axis.")
```

\begin{itemize}
\item How much variation is explained by the first two axes (PCs) of the ordination? Would you consider this an acceptable amount?
\item[] \textsf{The first two axes (PC1 and PC2) cumulatively explain  89\% of the variation, well beyond the minimum acceptable 70\%, so yes, this seems a good ordination.   \\ }
\item Which variable(s) contribute the most to PC1 and PC2? Along PC1, do you suspect they are negatively or positively correlated? How can you tell?
\item[] \textsf{\texttt{disp} and \texttt{mpg} contribute the most to PC1, while \texttt{qsec} and \texttt{drat} contribute the most to PC2. Along PC1, \texttt{disp} and \texttt{mpg} have a negative relationship because their loadings are negative and positive, respectively. \\ }
\end{itemize}


\pagebreak 

## Plot the PCA

```{r PCA, fig.cap='Three ways to plot the PCA. Note the use of \\texttt{vegan::make.cepnames} to automatically abbreviate the \\texttt{make.model} vector.', fig.subcap=c('Plain old base \\texttt{plot}.', 'With \\texttt{vegan::biplot} using \\texttt{base} graphics.', 'Here with \\texttt{ggplot}; calling the \\texttt{u} and \\texttt{v} objects require \\texttt{tibble::as\\_tibble} to properly format the objects.'), fig.width=3.5, fig.height=2.75, fig.ncol = 2}
# a
par(mar=c(4,5,0,2)) 
    plot(cars.pca, las=1)
# b
par(mar=c(4,5,0,2)) 
    biplot(cars.pca, display="species", las=1,
           xlim=c(-2,2), ylim=c(-1.75,1.5))
    text(cars.pca, display = "sites", 
         labels=make.cepnames(mtcars2$make.model))
# c    
    ggplot() + theme_ord(14) +
      labs(x="PC 1", y="PC 2") + 
      geom_vline(xintercept = 0, lty=2, color="darkgrey") +
      geom_hline(yintercept = 0, lty=2, color="darkgrey") +
      geom_segment(data=as_tibble(cars.pca$CA$v), 
                   aes(x=0, y=0, xend=PC1*0.90, yend=PC2*0.90), 
                   color="darkred", 
                   arrow = arrow(length = unit(0.02, "npc")))+
      geom_text(data=as_tibble(cars.pca$CA$v), 
                aes(x=PC1, y=PC2, 
                    label=row.names(cars.pca$CA$v)), 
                fontface="italic", color="darkred")  +
      geom_text(data=as_tibble(cars.pca$CA$u), 
                aes(x=PC1, y=PC2,
                    label=make.cepnames(mtcars2$make.model))) 
```

\clearpage

\begin{itemize}
\item True or False: The site scores are a mess.
\item[] \textsf{True! \\}
\item What might be done to improve this?
\item[] \textsf{A plot that excludes less-important or overlapping site scores would be easier to read. \\}
\end{itemize}

\textsf{\textbf{One option}: The \texttt{vegan} tutorial provides some solutions to plot labels that give priority to more-important rows (Fig. \ref{veganord}). \\}

\textsf{\textbf{Another option}: Using \texttt{geom\_label} instead of \texttt{geom\_text} helps with the readability of overlapping labels (Fig. \ref{ggord}). \\ }

```{r fig.cap='A couple of options for restricting the number of car types plotted. ', fig.subcap=c('Using some tricks from the vegan tutorial. \\label{veganord}', '\\texttt{ggplot} solution using \\texttt{geom\\_label} to help readability of overlapping labels. \\label{ggord}'), fig.height=4, fig.width=6, fig.ncol = 1}
# (a) Modified from the vegan tutorial
shnam <- make.cepnames(mtcars2$make.model)
 stems <- rowSums(cars.d) 
par(mar=c(4,5,0,2)) 
plot(cars.pca, dis="sites", type="n", las=1) 
ordilabel(cars.pca, dis="sites", 
          lab=shnam, priority = stems)
text(cars.pca, dis="sp", col="red")
# (b) Using geom_label instead of geom_text
    ggplot() + theme_ord(14) +
      labs(x="PC 1", y="PC 2") + 
      geom_vline(xintercept = 0, lty=3, color="darkgrey") +
      geom_hline(yintercept = 0, lty=3, color="darkgrey") +
      geom_segment(data=as_tibble(cars.pca$CA$v), 
                   aes(x=0, y=0, xend=PC1*0.90, yend=PC2*0.90), 
                   color="darkred", 
                   arrow = arrow(length = unit(0.02, "npc")))+
      geom_text(data=as_tibble(cars.pca$CA$v), 
                aes(x=PC1, y=PC2, 
                    label=row.names(cars.pca$CA$v)), 
                fontface="italic", color="darkred")  +
      geom_label(data=as_tibble(cars.pca$CA$u), 
                aes(x=PC1, y=PC2,
                    label=make.cepnames(mtcars2$make.model))) 
```

\pagebreak

## Environmental variables

### Plot 

```{r fig.cap='Two ways to show groups in the PCA. \\label{spiders}', fig.subcap=c('Plain old base \\texttt{plot}.', '\\texttt{ggplot} on the \\texttt{gg\\_ordiplot} object.'), fig.width=6, fig.height= 4, fig.ncol = 1}
# (a) 
par(mar=c(4,5,0,2))
plot(cars.pca, type="n", las=1, 
     xlim=c(-1.5,1.5))
ordispider(cars.pca, groups=mtcars2$continent, label=T)
# (b)
# Problem: gg_ordiplot looks nice enough, but can't be customized
# Solution: HACK IT
cont.gg <- gg_ordiplot(cars.pca, groups = mtcars2$continent, 
            spiders=TRUE, ellipse=FALSE, plot=FALSE) 

cont.gg$df_spiders %>% 
  plyr::rename(c("x"="X", 
                 "y"="Y") )%>%
  ggplot() + theme_ord(12) +
  geom_vline(xintercept = 0, lty=3, color="darkgrey") +
  geom_hline(yintercept = 0, lty=3, color="darkgrey") +
  labs(x="PC 1", y="PC 2") +
  geom_segment(aes(x=cntr.x, y=cntr.y,
                   xend=X, yend=Y, color=Group), 
               size=1.2, show.legend = FALSE) +
  geom_label(aes(x=cntr.x, y=cntr.y, 
                 label=Group, color=Group), 
             fontface="bold", size=4,
             label.size = 0, 
             label.r = unit(0.5, "lines"), 
             show.legend = FALSE)
```

\pagebreak 

### Test 

```{r}
f.fit <- envfit(cars.pca ~ continent, mtcars2)$factors
data.frame(Term=names(f.fit$r), 
           R.squared=round(f.fit$r, 2),
           P=f.fit$pvals, 
           row.names = NULL)   %>%
        pander("Results of envfit testing clusters by Continent in the PCA.")

tidy(pairwise.factorfit(cars.pca, mtcars2$continent, 
                   nperm = 999, p.method = "fdr") ) %>%
  unite(comparison, c("group1", "group2"), sep = " vs. ", remove=TRUE) %>%
    pander("Results of post-hoc pairwise comparison of Continent clusters in the PCA.") 
```

\textsf{Cars group by continent of origin, with significant differences between North American cars and those from both Asia and Europe. Most of this variability lies along the first axis (PC 1). }
\newpage 

Here's the script used to make \texttt{theme\_ord}: 


```{r eval=FALSE}
# Defining custom theme options
theme_ord <- function (base_size = 12, base_family = "") 
{
  theme_grey(base_size = base_size, base_family = base_family) %+replace% 
    theme(axis.text = element_text(size = rel(0.9)), 
          axis.title = element_text(face="bold"),
          axis.ticks = element_line(colour = "black"), 
          strip.text = element_text(face="bold"),
          legend.key = element_rect(colour = "grey80"), 
          panel.background = element_rect(fill = "white", colour = NA), 
          panel.border = element_rect(fill = NA, colour = "grey50"), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          plot.margin = unit(c(2,2,2,2), "mm"), 
          strip.background = element_rect(fill = "lightgreen", 
                                          colour = "grey50",
                                          size = 0.2)) 
}
```