---
title: "Plotting and NHST"
student:
  name: 
subtitle: 'Take-home quiz #1'
output:
  word_document: default
---
```{r echo=FALSE, results='hide'}
# Code Chunk 1
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
```{r echo=TRUE, results='hide'}
# Code Chunk 2
if (!require("pacman")) install.packages("pacman")
pacman::p_load(s20x)
```


# Getting started 

## Identify and prepare data

```{r echo=TRUE, eval=TRUE}
##
## Un-comment the dataset you want to use: 
##
  # data("attitude")
  # data("ToothGrowth")
  # data("mtcars")
```

## Load required packages 

```{r echo=TRUE}
pacman::p_load( )

```


# Distribution of response variable

## Plot

```{r echo=TRUE}
# Plot data distribution in a way that allows you 
# to evaluate which data model might fit best.

```

## Data model

* What is a Probability Distribution Function (PDF)?
* Which PDF seems appropriate for these data?

How well do the data fit the PDF you selected? Provide a graph here: 

```{r echo=TRUE}
# Plot data distribution in a way that allows you 
# to determine if data fit model assumptions.


```

# Data visualization 

```{r echo=TRUE, fig.width=6}
# Boxplot script here:



```

# Model fitting

## Which statistical model?

**Text answers:** 

* Which general statistical approach (type of test) might you apply to these data? 
* State a null and alternative hypothesis. 

### Run a test


```{r echo=TRUE}
# Provide script for statistical test here

```

### Post-hoc model validation 


```{r echo=TRUE}
# Provide script for post-hoc model validation here

```

### Post-hoc comparisons

```{r echo=TRUE}
# Provide script for post-hoc group comparison here

```

### Conclusions 

**Text answer:** 

Briefly describe the relationship between the variables in your dataset based on the results of the statistical test(s) conducted so far.

## Another statistical test?

### Additional data visualiation

```{r echo=TRUE}
# Means, SE, and line graph here

```

**Text answers:** 

* What does the graph signify?
* What changes, if any, does the graph prompt you to make to your statistical model? 

Depending on your second answer, as necessary: 

* Fit and present a new statistical model 
* Conduct and summarize a post-hoc test of model assumptions 
* Conduct post-hoc group comparisons, if needed. 

Finally, provide an updated conclusion about these data. 
