---
title: "Ecological ordination"
subtitle: "Homework Week 11"
author: "The Solution"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output: pdf_document
header-includes:
  - \usepackage{subfig}
  - \usepackage{multirow}
---

```{r setup, eval=TRUE, echo=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=FALSE, eval=TRUE, fig.align = 'center')
if (!require("pacman")) install.packages("pacman")
source('https://raw.githubusercontent.com/devanmcg/rangeR/master/R/CustomGGplotThemes.R')
source("https://raw.githubusercontent.com/devanmcg/rangeR/master/R/CustomCommunityFunctions.R")
pacman::p_load(vegan, tidyverse, broom, ggordiplots, RVAideMemoire, xtable )
```


```{r data}
 waders <- read_csv("C:/Users/Devan.McGranahan/GoogleDrive/Teaching/Classes/Analysis of Ecosystems/compiled notes/AoE course materials/rangeR/example datasets/wader_data.csv")
```

```{r unconstrained_ordination, results='hide'}
 wader.comm <- waders %>%
    select(., .data$Oystercatcher :.data$BlackWinged.Stilt )
  wader.mds <- capscale(wader.comm ~ 1, distance="bray", metaMDSdist=TRUE, 
                           engine=monoMDS, autotransform=TRUE) 

```

Which function/distance measure did you choose, and why?

\textsf{This ordination was fit with the \texttt{capscale} function using the Bray-Curtis distance measure and the \texttt{metaMDS} options because this fit these data best and had the maximal variance explained in the top two axes. \\}

How can you be confident your ordination is a decent representation of variation in these data? 

\textsf{More than 70\% of variation is explained by the first two axes (Table \ref{VarView}), after which further inertia is marginal (Fig. \ref{scree}). \\}


```{r assessment, results='asis'}
var.view(wader.mds, 5) %>%
  xtable("Proportion variance explained by the first five axes of the MDS.", label="VarView") %>%
          print(comment=FALSE, include.rownames=TRUE)
```

```{r screeplot, results='hide', fig.cap='Screeplot for the Bray-Curtis MDS on the wader data. \\label{scree}', fig.width=7, fig.height=3}

enframe(wader.mds$CA$eig) %>%
  plyr::rename(c("value"="Inertia")) %>%
    slice(1:9) %>%
ggplot() + theme_bw(14) + 
  geom_line(aes(x=name, y=Inertia, group=1), 
            color="darkblue", size=1.5) + 
  labs(x="Axis")
```


Which categorical variable(s) show the most difference in the bird communities? Which show the least? Provide evidence/examples.

\textsf{As the only factor with \emph{P} > 0.10, Location provides the least explanation of difference in the bird communities, while Habitat is the best (Table \ref{fit}). \\}

```{r envfit, results='asis'}
w.fit <- envfit(wader.mds ~ Country + Ocean + Location + Habitat, data=waders)$factors
```

```{r fit_tab, results='asis'}
tibble(Term=names(w.fit$r), 
           `R squared`=round(w.fit$r, 2),
           `P value`=w.fit$pvals) %>%
      xtable("Results of \\texttt{factorfit} on wader data. ", label="fit") %>%
          print(comment=FALSE, include.rownames=FALSE)
```

Which bird species contribute the most to variation among habitat types?

\textsf{ This is a case where a constrained ordination most directly answers the question. Fitting \texttt{Habitat} as a constraint, Whitefronted Plover and Avocet are at the opposite ends of the first (constrained) axis (-0.73 and 0.44, respectively; Table \ref{cap}). \\}

```{r constrained_ordination, results='hide'}
hab.cap <- capscale(wader.comm ~ waders$Habitat, distance="bray", metaMDSdist=TRUE, 
                           engine=monoMDS, autotransform=TRUE)
```

```{r spp_scores, results='asis'}
scores(hab.cap, display="species") %>%
            as.data.frame() %>%
              rownames_to_column("species") %>%
                select(-MDS1) %>%
                  arrange(desc(abs(CAP1))) %>%
        xtable("Species ranked by absolute value of species score along \\texttt{CAP1} in the constrained \\texttt{capscale} model $[y_n] \\sim\\ Habitat$. \\texttt{CAP1} accounts for 75\\% of variance in the data. The constrained model is significantly better than the unconstrained ordination (Table \\ref{comp}). ", label="cap") %>%
          print(comment=FALSE, include.rownames=FALSE)
```

```{r comp, results='asis'}
anova(wader.mds, hab.cap) %>%
  xtable("Results of statistical comparison of constrained vs. unconstrained models.", label="comp") -> comp

addtorow <- list()
addtorow$pos <- list(-1)
addtorow$command <- 
  paste0(
  paste0('\\multicolumn{5}{l}{Model 1: wader.comm $\\sim$ 1 } \\\\', collapse=''), 
  paste0('\\multicolumn{5}{l}{Model 2: wader.comm $\\sim$ Habitat}\\\\ ', collapse=''))

print(comp, add.to.row=addtorow, include.colnames=T, 
      comment=FALSE, include.rownames=FALSE)
```




Which bird species seems most abundant along Northern Namibian coastlines?

\textsf{Turnstones are the species most associated with Namibian coastlines (Fig. \ref{ord}).  \\}

```{r MDS, fig.cap='Species scores and site scores plotted by habitat and country (N=Namibia, SA=South Africa). \\label{ord}', fig.width=7, fig.height=4}
spp.sc <- scores(wader.mds, display="species") %>%
            as.data.frame() %>%
              rownames_to_column("species")

site.sc <- scores(wader.mds, display="sites") %>%
            as.data.frame() %>%
              mutate(Country = ifelse(waders$Country == "Namibia", "N", "SA"), 
                     Habitat = waders$Habitat)

  ggplot() + theme_ord(12) +
  geom_vline(xintercept = 0, lty=3, color="darkgrey") +
  geom_hline(yintercept = 0, lty=3, color="darkgrey") +
  labs(x="MDS 1", y="MDS 2") +
  geom_text(data=filter(spp.sc, abs(MDS1) > 0.3), 
            aes(x=MDS1*1.0, y=MDS2*1.5, 
                label=species), 
            fontface="italic", 
            check_overlap = TRUE) +
  geom_text(data=site.sc, aes(x=MDS1, y=MDS2, 
                               label=Country, 
                              color=Habitat), 
            fontface="bold", 
            check_overlap = TRUE) + 
  scale_color_manual(name="Habitat", values = cbPal5) 
```


\clearpage 


# Appendix
## Full session script
```{r ref.label=knitr::all_labels(!label %in% c('fit_tab', 'comp', 'sessionInfo')),echo=TRUE,eval=FALSE}
```

## Session info
```{r sessionInfo, echo=FALSE}
sessionInfo();
```