---
title: "Spatial interpolation"
subtitle: "Homework Week 12"
author: "The Solution"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output: pdf_document
header-includes:
  - \usepackage{subfig}
---

```{r setup, eval=TRUE, echo=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=FALSE, eval=TRUE, fig.align = 'center')
if (!require("pacman")) install.packages("pacman")
source('https://raw.githubusercontent.com/devanmcg/rangeR/master/R/CustomGGplotThemes.R')
pacman::p_load(plyr, tidyverse, ggmap, akima, viridis, xtable)
#pacman::p_load_gh("dkahle/ggmap")

```


```{r data}
#
# Load windbreak data from github
#
D = url("https://github.com/devanmcg/rangeR/blob/master/example%20datasets/windbreak.data.Rdata?raw=true")
load(D)

wbd <- merge(windbreak.data$windbreak.points, 
             windbreak.data$windbreak.densities, 
             by = "id") %>%
          mutate(change = density.2015 - density.2010)
#
# Get North Dakota boundaries
#
  NDs <- map_data("state") %>% filter(region =="north dakota")
  NDc <- map_data("county") %>% filter(region =="north dakota")
```


What is the spatial pattern of change in windbreak density between 2010 and 2015?

\textsf{Windbreak density declined in a few discrete pockets, primarily where North Dakotans would call "east river," east of and, geologically similar but geographically illogical, north of the Missouri River (Fig. \ref{change}). \\}

```{r change_interp, fig.cap="Spatial interpolations of windbreak data fit with \\texttt{akima} package.", fig.subcap=c('Spatial pattern of windbreak removal in North Dakota between 2010 and 2015. \\label{change}', 'Windbreak density across North Dakota as of 2015. \\label{density}'), fig.height=3, fig.ncol = 1}
#
# Interpolating change
#
 ChangeInterp <-  with(wbd, interp(x=long, y=lat, 
                                        z=change))
ChangeInterp.df <- as.data.frame(interp2xyz(ChangeInterp))
names(ChangeInterp.df) <- c("long", "lat", "change")
#
# Plotting change 
#
 ggplot()  + coord_map("polyconic") + theme_map() +
    geom_tile(data=ChangeInterp.df, aes(x=long, y=lat, 
                                fill=change)) + 
         geom_path(data=NDc, 
                aes(x=long, y=lat, group=group), 
                color="grey60", size=0.5) + 
      geom_path(data=NDs, 
                aes(x=long, y=lat, group=group), 
               color="grey30", size=0.75) +
    scale_fill_viridis(name="Change in\nwindbreak\ndensity\n2010-2015", 
                       na.value="transparent", 
                       direction = -1) 
#
# Interpolating density
#
DenseInterp <-  with(wbd, interp(x=long, y=lat, 
                                        z=density.2015))
DenseInterp.df <- as.data.frame(interp2xyz(DenseInterp))
names(DenseInterp.df) <- c("long", "lat", "density")
#
# Plotting density 
#
 ggplot()  + coord_map("polyconic") + theme_map() +
    geom_tile(data=DenseInterp.df, aes(x=long, y=lat, 
                                fill=density)) + 
         geom_path(data=NDc, 
                aes(x=long, y=lat, group=group), 
                color="grey60", size=0.5) + 
      geom_path(data=NDs, 
                aes(x=long, y=lat, group=group), 
               color="grey30", size=0.75) +
    scale_fill_viridis(name="Windbreak\ndensity\n(2015)", 
                       na.value="transparent", 
                       direction = 1) 

```


What parts of the state, geographically and by ecoregion, have the most windbreaks? Which have the least?

\textsf{Geographically, the eastern part of the state has the most windbreaks, while the southwestern corner has the least (excluding a few localized pockets; Fig. \ref{density}).  \\
By ecoregion, the Lake Agassiz Plain (basically the Red River Valley) had the highest windbreak density, while the NW Plains had the lowest (Table \ref{densetab}). \\}

```{r ecoregion_density, results='asis'}
#
# Ecoregion summary tables
#
wbd  %>%
  group_by(ecoregion) %>%
    summarize(`mean density` = round(mean(density.2015), 0)) %>%
      arrange(desc(`mean density`)) %>%
      xtable("Average density of windbreaks in North Dakota in 2015, by ecoregion.",
             label="densetab") %>%
          print(comment=FALSE, include.rownames=FALSE)
```

Which parts of the state, geographically and by ecoregion, showed the greatest change in windbreak density during the study region? Which showed the least?

\textsf{As shown in Fig. \ref{change}, the eastern part of the state, especially the Red River Valley, had the greatest reductions, while the southwestern part of the state appears unchanged.  \\ 
By ecoregion, the Lake Agassiz Plain had the greatest reduction, while the NW Great Plains was entirely unchanged between 2010-2015 (Table \ref{changetab}). \\ }

```{r ecoregion_change, results='asis'}
wbd  %>%
  group_by(ecoregion) %>%
    summarize(`mean change` = round(mean(change), 0)) %>%
      arrange(`mean change`) %>%
      xtable("Average of change in windbreak density between 2010 and 2015 in North Dakota, by ecoregion.", 
            label="changetab") %>%
          print(comment=FALSE, include.rownames=FALSE)
```



\clearpage 


# Appendix
## Full session script
```{r ref.label=knitr::all_labels(!label %in% c('fit_tab', 'comp', 'sessionInfo')),echo=TRUE,eval=FALSE}
```

## Session info
```{r sessionInfo, echo=FALSE}
sessionInfo();
```