---
title: "Geolocations with geonames package for R"
subtitle: "A workflow for using ropensci to access the free geonames API"
author: "DAM"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    keep_md: yes
---

```{r eval=TRUE, message=FALSE, echo=FALSE}
# Code Chunk 1
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=TRUE, eval=TRUE, fig.align = 'center')
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse)
options(geonamesUsername="devanmcg")
```

# Introduction 

Since **ggmap**'s *geocode* function stopped accessing Google Map's API like it used to, we turn to other solutions to getting coordinates for place names. Having played with a couple I've landed on the **geonames** package from **ropensci**, which creates a connection between **R** and the open data source geonames. Main benefits include: 

* It is free (although a free account is required to use)
* It has a lot of options for specifying searches 

Compared to *geocode* this last point makes it a little less straightforward to use, but overall I find it better. 

# Get set up 

1. Create a free geonames account at http://www.geonames.org/login/. 

2. Install the package from github using the **devtools** package:

```{r eval=FALSE} 
devtools::install_github("ropensci/geonames")
```

3. Set your **R** environment's global options to include your username for geonames:

```{r eval=FALSE}
options(geonamesUsername="USERNAME")
```

4. Load the library
```{r }
library(geonames)
```

# Using the package

The key function here is *GNsearch*, which searches the geonames database online. Let's start by just searching for "Fargo:"

```{r eval=FALSE}
GNsearch(name_equals = "Fargo")
```

We see in the output below that the database has more places named "Fargo" than we might have anticipated... 19 in fact, including a stream in Djibouti and a farm in Norway. Scroll down below the table and see how we refine the search. 

```{r echo=FALSE}
GNsearch(name_equals = "Fargo") %>%
   pander::pander("All the places returned by searching geonames for 'Fargo'." ) 
```

## Refining *GNsearch* results

You should notice immediately that along with latitude and longitude, the results include a lot of information about the various places under some unclear column names. We can use this information to narrow our search results. An overview of each of the parameters -- column names -- [is available here](http://www.geonames.org/export/geonames-search.html). 

Not always do column/parameter names match up with arguments to the *GNsearch* function. For example, *featureClass* is the argument that maps to the *fcl* parameter. For many uses this is probably the finest search parameter one might use: P = populated place, L = parks, etc. Much much finer designations can be given with the *fcode* or *fcodeName* parameter ([see here](http://www.geonames.org/export/codes.html) ). For example, Fargo, ND and Fargo, AR are both P for populated places, but only Fargo, ND is fcode PPLA2, for "seat of a second-order administrative division", or county seat. (States are first-order administrative divisions).

Because we know we want the Fargo in the state of North Dakota, we specify adminCode1="ND" and only one result is returned: 

```{r eval=FALSE}
GNsearch(name_equals = "Fargo", adminCode1="ND")
```

```{r echo=FALSE}
GNsearch(name_equals = "Fargo", adminCode1="ND") %>%
   pander::pander("All the places returned by searching geonames for 'Fargo' in the state of North Dakota." ) 
```

## Cut to the chase

Here's the script used to generate the list of towns and RECs we used in class [nd.features.d](https://drive.google.com/open?id=1BwVWWHictwrYEPg5MVssG0igVyVMJnkh): 
```{r }

nd.features <- data.frame(feature=c("city", "city", "city", "city", 
                                    "rec", "rec", "rec", "rec", 
                                    "rec", "rec"), 
                          name=c("Fargo", "Grand Forks", "Bismarck", "Minot", 
                                 "Hettinger", "Williston", "Carrington","Dickinson",
                                 "Streeter", "Langdon") )
nd.features.d <- 
  nd.features %>%
    split(.$name) %>%
      map( ~ GNsearch(name_equals = .x$name, country = "US", 
                      adminCode1="ND", featureClass="P")) %>%
        map_dfr(~ (.))  %>%
    full_join(nd.features) %>%
    mutate(state = adminName1, 
           long=as.numeric(lng), 
           lat = as.numeric(lat)) %>%
    select(feature, name, state, long, lat) 
    
```

```{r echo=FALSE}
pander::pander(nd.features.d, "Longitude and latitude of cities and RECs in North Dakota")
```

```{r}
NDs <- map_data("state") %>%
  filter(region == "north dakota")

NDc <- map_data("county") %>%
  filter(region == "north dakota")

ggplot() + theme_bw() + 
  geom_path(data=NDc, 
            aes(x=long, y=lat, group=group), 
            color="dark grey") +
  geom_path(data=NDs, 
            aes(x=long, y=lat, group=group), 
            color="black") +
  geom_point(data=nd.features.d, 
             aes(x=long, y=lat, 
                shape=feature, color=feature)) +
  geom_text(data=nd.features.d, 
            aes(x=long, y=lat, 
                label=name), nudge_y = 0.15) +
  coord_map("conic", 60) 

```

