---
title: "Analysis of Ecosystems homework week 6"
subtitle: "Multiple regression and ANOVA"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "Put your name here"
output:
  word_document: default
---

```{r echo=FALSE, results='hide'}
# Code Chunk 1
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```
```{r echo=TRUE, results='hide'}
# Code Chunk 2
if (!require("pacman")) install.packages("pacman")
pacman::p_load(s20x, plyr, ggplot2, car, gvlma)
```

# Data preparation

## Identification

```{r echo=TRUE, results='verbatim'}
# Code Chunk 3: Data structure




```


## Assumptions

```{r echo=TRUE, results='verbatim', fig=TRUE}
# Code Chunk 4: Model assumptions




```


# Multiple linear regression 

## Graphing

```{r echo=TRUE, results='verbatim', fig=TRUE}
# Code Chunk 5: Continuous vs Continuous and Categorical graph




```


## Hypothesis statement

Null hypothesis: 
Alternative hypothesis that refers to two variables simultaneously:

## Fit a linear model

```{r echo=TRUE, results='verbatim'}
# Code Chunk 6: Fit linear model




```


## Interpret

Summarize the results and present a conclusion with specific reference to the above hypotheses:

# ANOVA on categorical predictor variables 

## Graphing

```{r echo=TRUE, results='verbatim', fig=TRUE}
# Code Chunk 7: Continous vs. Categorical graph




```


## Hypothesis statement

Null hypothesis:
Alternative hypothesis:

## Fit an ANOVA model

```{r echo=TRUE, results='verbatim'}
# Code Chunk 8: ANOVA mdoel




```
 

## Interpret

Summarize the results and present a conclusion with specific reference to the hypotheses you stated above. 