---
title: "Analysis of Ecosystems homework week 6"
subtitle: "Multiple regression and ANOVA"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "The solution"
header-includes:
   - \usepackage{nicefrac}
output: pdf_document
---

```{r chunk1, echo=FALSE, results='hide', message=FALSE}
# DOn't mess with this chunk
knitr::opts_chunk$set(message = FALSE, warning=FALSE, echo=TRUE, eval=TRUE, fig.align = 'center')
if (!require("pacman")) install.packages("pacman")
```

```{r chunk2, echo=TRUE, results='hide'}
pacman::p_load(s20x, pander, plyr, tidyverse, gridExtra, multcomp)
```

# Data preparation

## Identification

```{r }
mpg <- mpg %>%
  mutate(trans = ifelse((substring(trans,1,4)=='auto'),'auto', 'manual'), 
          drv = case_when(
           drv == "f" ~ "front-wheel", 
           drv == "4" ~ "four-wheel", 
           TRUE ~ "rear-wheel"    )) %>%
  mutate(drv = as.factor(drv)  )
str(mpg)
```

\textsf{\texttt{cty} will be the response variable in all tests. Multiple regression will test \texttt{cty} against \texttt{displ} as a continuous variable and \texttt{trans} as a categorical variable. ANOVA will test \texttt{cty} against \texttt{drv}. }

## Assumptions

\textsf{Before fitting a model, one needs to ensure that the distribution of the response variable fits a normal distribution:}

```{r chunk4, fig.height = 6, fig.cap="TOP: Distribution of the \\texttt{cty} variable, before log transformation (L) and after log transformation (R). BOTTOM: Q-Q plots for the above. \\label{dist}"}
# Distribution plots
  raw.d.gg <-
    ggplot(mpg, aes(x=cty)) + theme_bw(14) + 
      geom_histogram(aes(y=..density..),      
                     binwidth=1,
                     colour="black", fill="lightgreen") +
      geom_density(alpha=0.2, fill="lightgreen") +
      stat_function(data=mpg, 
                          fun = dnorm, 
                          args=list(mean=mean(mpg$cty),    
                                      sd=sd(mpg$cty)),    
                          colour="blue", size=1.1) 
  log.d.gg <-
    ggplot(mpg, aes(x=log(cty))) + theme_bw(16) + 
      geom_histogram(aes(y=..density..),      
                     binwidth=0.1,
                     colour="black", fill="lightgreen") +
      geom_density(alpha=0.2, fill="lightgreen") +
      stat_function(data=mpg, 
                          fun = dnorm, 
                          args=list(mean=mean(log(mpg$cty)),    
                                      sd=sd(log(mpg$cty))),    
                          colour="blue", size=1.1) 
# Q-Q plots 
  raw.QQ.gg <- 
         ggplot(mpg, aes(sample=cty)) + theme_bw(14) +
            stat_qq(size=4,  bg="#43a2ca", col="black", pch=21) +
            stat_qq_line(size=1.5, color="blue")
  log.QQ.gg <- 
         ggplot(mpg, aes(sample=log(cty))) + theme_bw(14) +
            stat_qq(size=4, bg="#43a2ca", col="black", pch=21) +
            stat_qq_line(size=1.5, color="blue")

grid.arrange(raw.d.gg,  log.d.gg, 
             raw.QQ.gg, log.QQ.gg, nrow = 2)
```


\textsf{The untransformed data are skewed right (Fig. \ref{dist}), but log transformation improves the fit between the distribution of the \texttt{cty} variable and the theoretical normal distribution. QQ plot of the log transformed data confirm the better fit. }

 

# Multiple linear regression 

## Graphing

```{r MR, out.width='60%', fig.cap="The linear relationship between city fuel economy (\\texttt{cty}) against engine size (\\texttt{displ}) and \\texttt{trans}mission type. \\texttt{cty} log transformed. \\label{lm}"}
prov.col2 <- viridis::magma(n=2, begin=0.25, end=0.75, direction = 1) 
ggplot(mpg, aes(x=displ, y=log(cty))) + theme_bw(20) + 
  geom_smooth(aes(color=trans), 
              se=FALSE, method="lm", 
              size=1.5) + 
  geom_point(aes(fill=trans, shape=trans), 
             size=4,  col="black") +
  scale_shape_manual(name="Transmission\ntype", 
                     values = c(21, 24)) + 
  scale_fill_manual(name="Transmission\ntype", 
                     values = prov.col2) + 
  scale_color_manual(name="Transmission\ntype", 
                     values = prov.col2) +
  labs(x="Engine size (L)", 
       y="City fuel economy\n(log mpg)") 
```


## Hypothesis statement

\begin{itemize}
\item[] \textsf{$H_0$: No relationship between city fuel economy and transmission type or engine size.}
\item[] \textsf{$H_1$: City fuel economy declines with engine size; manual transmissions generally get better fuel economy at a given engine size.}
\end{itemize}

## Fit a linear model

```{r}
m1 <- lm(log(cty) ~ displ + trans, mpg)
car::Anova(m1, type="2")
summary(m1)
gvlma::gvlma(m1)
```


## Interpretation

\textsf{The model fit these data well, with approximately 68\% of variance explained. Results of \texttt{gvlma} analysis of the multiple regression model show that assumptions for both skewness and heteroscedasticity were met, indicating that the model was robust. Thus, we can reject that null hypothesis, and have statistical evidence for both parts of the alternative hypothesis: both terms were significant, and the \emph{t} statistics for \texttt{displ} and manual \texttt{trans}missions were -20 and 2, respectively. }

# ANOVA on categorical predictor variables 

## Graphing

```{r ANOVA, out.width='60%', fig.cap="The relationship between city fuel economy (\\texttt{cty})  and drive train type. \\texttt{cty} log transformed in analysis, but not in the graph. \\label{anova}"}

prov.col3 <- viridis::magma(n=3, begin=0.25, end=0.75, direction = 1) 

ggplot(mpg, aes(x=drv, y=cty)) + theme_bw(20) + 
  geom_boxplot(aes(fill=drv), 
             size=1.5,  col="black") +
  scale_fill_manual(values = prov.col3, guide=FALSE) + 
  labs(x="Drive train", 
       y="City fuel economy\n(mpg)") 
```


## Hypothesis statement

\begin{itemize}
\item[] \textsf{$H_0$: No relationship between city fuel economy and drive train.}
\item[] \textsf{$H_1$: Relative to front wheel drive vehicles, city fuel economy will be lower for four wheel drive vehicles and lowest for rear wheel drive vehicles.}
\end{itemize}

## Fit an ANOVA model

```{r }

m2 <- lm(log(cty) ~ drv, mpg)
anova(m2) 
m2.tuk <- glht(m2, linfct = mcp(drv="Tukey"))
summary(m2.tuk)

```
 

## Interpretation

\textsf{As expected, there was significant variability in city fuel economy among the drive train types (Fig. \ref{anova}): front wheel drive vehicles were consistently greater than both rear wheel drive (\emph{t} = 13.2, \emph{P} < 0.001) and four wheel drive vehicles (\emph{t} = 8.5, \emph{P} < 0.001). However, there was no statistically-significant difference between rear wheel drive and four wheel drive vehicles (\emph{t} = -0.24, \emph{P} = 0.97). }