---
title: "Model selection and confidence intervals"
subtitle: 'Homework week 7'
author: "The Solution"
output:
  pdf_document: default
---
```{r echo=FALSE, message= FALSE, results='hide'}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=TRUE, eval=TRUE, 
                      fig.align = 'center')
if (!require("pacman")) install.packages("pacman")

```

```{r}
pacman::p_load(s20x, tidyverse, gvlma, AICcmodavg, gridExtra, xtable)
```

# Data preparation

## Identification

```{r }
data(attitude)
str(attitude)
```

## Assumptions

```{r fig.width=8, fig.cap="Distribution and Q-Q plot for response variable \\texttt{rating}. \\label{dist}"}
rat.dist.gg <- 
  attitude %>%
    ggplot(aes(x=rating)) + theme_bw(16) + 
      geom_density(alpha=.2, fill="#FF6666") +
      geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", 
                 fill="lightgreen") +
      labs(x="rating") +
      geom_line(data=data.frame(
                        X=seq(25, 100, 1), 
                        Y=dnorm(x=seq(25, 100, 1),
                                mean=mean(attitude$rating),    
                                sd=sd(attitude$rating))), 
                aes(x=X, y=Y), 
                    colour="blue", size=1.1)
rat.QQ.gg <-
  attitude %>%
    ggplot(aes(sample=rating)) + theme_bw(16) +
            stat_qq(size=4, bg="#43a2ca",
                    col="black", pch=21) +
            stat_qq_line(size=1.5, color="blue")
grid.arrange(rat.dist.gg, rat.QQ.gg, nrow = 1)
```

\textsf{The data are sufficiently symetrical around the mean that a normal (Gaussian) distribution fits them well. }

```{r results='hide'}
mod.sum <- summary(gvlma(lm(rating ~ complaints + privileges + learning,
                        data=attitude )))
```

```{r results='asis'}
xtable(mod.sum, caption="The full potential model meets assumptions of the linear model.",
       label="tab:gvlma") %>%
  print(comment=FALSE)
```

# Model fitting and selection 

```{r echo=FALSE, eval=FALSE, out.width='80%', fig.cap ="Scatterplot matrix of \\texttt{rating} against potential predictor variables. \\label{spm}"}
attitude %>%
  dplyr::select(rating, complaints, privileges, learning) %>%
GGally::ggpairs( progress = FALSE) + 
  theme_bw(14) +
  theme(strip.text.y = element_text(angle = 0)) 

# \textsf{Low correlation coefficents (\emph{r} < 0.2; Fig. \ref{spm}) between \texttt{rating} and \texttt{critical} and \texttt{advance} suggest they should not be considered in potential models. }
```



```{r results='asis'}
car::vif(lm(rating ~ complaints + privileges + learning, data=attitude )) %>%
  t() %>%
  as.data.frame() %>%
  xtable(caption="Low Variable Inflation Factors for each potential
         predictor variables.") %>%
  print(comment=FALSE, include.rownames=FALSE)
```


## Define model set

```{r}
null <- lm(rating ~ 1, attitude) 
C <- lm(rating ~ complaints, attitude)
P <- lm(rating ~ privileges, attitude)
L <- lm(rating ~ learning, attitude)
CP <- lm(rating ~ complaints + privileges, attitude)
CL <- lm(rating ~ complaints + learning, attitude)
PL <- lm(rating ~ privileges + learning, attitude)
CPL <- lm(rating ~ complaints + privileges + learning, attitude)

cand.mod.names <- c("null", "C", "P", "L", "CP", "CL", "PL", "CPL")
```

## Model selection

```{r results='asis'}
cand.mods <- list( )
for(i in 1:length(cand.mod.names)) {
  cand.mods[[i]] <- get(cand.mod.names[i]) }
library(AICcmodavg)
aictab(cand.set = cand.mods,          
       modnames = cand.mod.names) %>%
  xtable(caption="Model rankings based on $AIC_{c}$ information criterion.
         Models identified by first letter of predictor variable names.\\label{MS}") %>%
    print(comment=FALSE, include.rownames=FALSE)
```



\textsf{All models that include \texttt{complaints} can be considered competitive in $AIC_{c}$-based model selection (Table \ref{MS}).  }

## Model averaging

```{r results='asis'}
terms <- c("(Intercept)", "complaints", "learning", "privileges")
        av.params <- as.data.frame(array(NA,c(length(terms),4)))
        colnames(av.params)<-c("term","estimate","ciL","ciU")
       for(i in 1:length(terms)) {
        av <- modavg(parm = paste(terms[i]), 
                     cand.set = cand.mods, 
                     modnames = cand.mod.names)
            av.params[i,1] <- terms[i]
            av.params[i,2] <- round(av$Mod.avg.beta, 2)
            av.params[i,3] <- round(av$Lower.CL, 3) 
            av.params[i,4] <- round(av$Upper.CL, 3) }
av.params %>%
  xtable(caption="Averaged regression coefficients on top-ranked models
         from $AIC_{c}$-based model selection (Table \\ref{MS}).
         \\label{MA}") %>%
      print(comment=FALSE, include.rownames=FALSE)
```

\textsf{As the slope term of the regression equation denotes the strength of the modelled relationship, the estimated coefficient for each slope term in a regression model can be interpreted as a measure of that term's relative importance to the response variable, or \emph{effect size}.}

## Plot confidence intervals

```{r fig.width=6, fig.cap="Model-averaged 95\\% confidence intervals with regression coefficient estimates for terms in top-ranked models (Table \\ref{MS}). \\label{CIs} "}
av.params %>%
  filter(terms != "(Intercept)") %>%
    ggplot() + theme_bw(16) +
      coord_flip() + 
      geom_hline(yintercept = 0) + 
      geom_errorbar(aes(x=term,
                        ymin=ciL, 
                        ymax=ciU), 
                        width=0.1, size=1, 
                    color="#377eb8") +
      geom_point(aes(x=term, 
                     y=estimate), 
                 size=4, pch=21, stroke=2, 
                 bg="#377eb8", color="white")
```

# Conclusions

\textsf{$AIC_{c}$-based model selection indicated that three variables, \texttt{privileges}, \texttt{learning}, and \texttt{complaints} were associated with employee \texttt{ratings} of job satisfaction (Table \ref{MS}). In comparing model-averaged regression coefficients (Fig. \ref{CIs}), only \texttt{complaints} had a non-zero relationship with \texttt{rating}, which was positive. 95\% CIs for \texttt{learning} and \texttt{privileges} overlapped zero; these terms had positive and negative trends with \texttt{rating}, respectively. Thus, employee ratings appear to be most strongly determined by how well eomployees felt the company handled employee complaints.} 